---
title: "How to work with ranges of HS-codes"
author: "Aleksandr Matrunich"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{How to work with ranges of HS-codes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = F}
options(width = 120) 
knitr::opts_chunk$set(echo = F)
suppressWarnings(suppressPackageStartupMessages(library(dplyr)))
library(stringr)
library(hsfclmap)
source(file.path(Sys.getenv("HOME"),
                 "r_adhoc",
                 "privateFAO",
                 "OrangeBook",
                 "tradeR",
                 "printTab.R"))

hsfclmap <- hsfclmap %>% 
  mutate(tocode = trailingZeros(fromcode, tocode))

```

The map on the HS-side consists of two columns which define range of HS-codes for matching to single FCL-code.

```{r results = "asis"}
hsfclmap %>% 
  sample_n(6) %>% 
  printTab(caption = "Example of links from MDB-files")
```


How many unique HS codes in fromcode field?
```{r results = "asis"}
hsfclmap %>% 
  select(fromcode) %>% 
  distinct() %>% 
  summarize(unique_codes = n()) %>% 
  printTab(caption = "In character format")

hsfclmap %>% 
  select(fromcode) %>% 
  mutate(fromcode = as.numeric(fromcode)) %>% 
  distinct() %>% 
  summarize(unique_codes = n()) %>% 
  printTab(caption = "In numeric format")

hsfclmap %>% 
  select(fromcode, tocode) %>% 
  filter(fromcode != tocode) %>% 
  select(fromcode) %>% 
  mutate(fromcode = stringr::str_trim(fromcode)) %>% 
  distinct() %>% 
  mutate(codenum = as.numeric(fromcode)) %>% 
  group_by(codenum) %>% 
  mutate(n = n()) %>% 
  ungroup() %>% 
  filter(n > 1) %>% 
  arrange(codenum) %>% 
  mutate(id = dense_rank(codenum)) %>% 
  select(id, fromcode, asnumeric = codenum) %>% 
  as.data.frame() %>% 
  printTab(caption = "HS codes what have non-unique numeric equivalent")

```

We could remove leading zeros, but how to detect differences between 02nd and 20th chapters?


1. Fromcode must be less or equal than tocode.
2. Tocode of previous row must be less than fromcode of current row.

```{r, results = "asis"}
hsfclmap %>% 
  select(fromcode, tocode) %>% 
  distinct() %>% 
  mutate(fromnum = as.numeric(fromcode),
         tonum   = as.numeric(tocode),
         fromlessto = fromnum <= tonum) %>% 
  filter(!fromlessto) %>% 
  select(-fromlessto) %>% 
  printTab(caption = "HS-codes where fromcode is less than tocode")
```

So we need to add a rule, if tocode is shoter than fromcode, add trailing zeros to tocode. But in some cases (fromcode = 29192 and tocode = 19192) it is a mistake.


```{r results = "asis"}
hsfclmap %>% 
  select(area, flow, validyear, fromcode, tocode, fcl) %>% 
  distinct() %>% 
  mutate(fromnum = as.numeric(fromcode),
         tonum   = as.numeric(tocode)) %>% 
  group_by(area, flow) %>% 
  arrange(fromnum) %>% 
  mutate(tolessnextfrom = tonum < lead(fromnum)) %>% 
  mutate(toinvestigate = !tolessnextfrom | 
           !lag(tolessnextfrom)) %>% 
  mutate(group_id = paste(area, flow, fromcode, tocode, sep = "-"),
         group_id = ifelse(!tolessnextfrom, group_id, lag(group_id))) %>% 
  filter(toinvestigate) %>% 
  group_by(group_id) %>% 
  mutate(fcl_div = length(unique(fcl)) > 1) %>% 
  ungroup() %>% 
  filter(fcl_div) %>% 
  mutate(group_id = as.numeric(as.factor(group_id)),
         id = dense_rank(group_id)) %>% 
  select(id, area, flow, fromcode, tocode, fcl) %>% 
  as.data.frame() %>% 
  printTab(caption = "Links where next fromcode is not more than previous tocode and there are different FCL codes")
  
```

# Trailing zeros

The following table shows links where fromcode and tocode have different length. We need to add trailing zeros.
```{r results = "asis"}
hsfclmap %>% 
    # select(fromcode, tocode) %>% 
  # distinct() %>% 
  filter(stringr::str_length(fromcode) != stringr::str_length(tocode)) %>% 
  printTab()
```


# Corrections

1. If tocode is shorter than fromcode, add to to code trailing zeros.
2. Leading zeros will be automatically removed during conversion to numeric type.

```{r}
# hsfclmap %>% 
  
```
